---
title: "Salary Data Analysis Dashboard (2020-2023)"
author: "Ahmad Ammar"
date: "2024-12-22"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
    theme: cosmo
    social: menu
    source_code: embed
runtime: shiny
---

<style>
.sidebar {
  background-color: #f8f9fa;
  padding: 20px;
  border-right: 3px solid #007bff;
}

.chart-title {
  color: #007bff;
  font-weight: bold;
  border-bottom: 2px solid #007bff;
  padding-bottom: 10px;
  margin-bottom: 20px;
}

.value-box {
  border-radius: 10px;
  box-shadow: 0 4px 6px rgba(0,0,0,0.1);
}

h1, h2, h3, h4 {
  font-family: 'Segoe UI', Tahoma, Geneva, Verdana, sans-serif;
}

.sidebar h3 {
  color: #495057;
  line-height: 1.6;
}

hr {
  border: 0;
  height: 2px;
  background: linear-gradient(to right, #007bff, #6c757d);
  margin: 20px 0;
}
</style>

```{r setup, include=FALSE}
library(flexdashboard)
library(tidyverse)
library(plotly)
library(DT)
library(shiny)
library(readxl)
library(dplyr)
library(janitor)
library(tidyr)
library(scales)
```

```{r data-loading, include=FALSE}
# Set file paths
data_path <- "C:/Users/HP/Desktop/R FINAL PROJECT/data/"

# Load data from Excel files
Data_2020_xlsx <- read_excel(paste0(data_path, "Data_2020.xlsx"))
Data_2021_xlsx <- read_excel(paste0(data_path, "Data_2021.xlsx"))
Data_2022_xlsx <- read_excel(paste0(data_path, "Data_2022.xlsx"))
Data_2023_xlsx <- read_excel(paste0(data_path, "Data_2023.xlsx"))

# Combine all years
concatenated_data <- rbind(Data_2020_xlsx, Data_2021_xlsx, 
                           Data_2022_xlsx, Data_2023_xlsx)

# Clean data
combined_data_clean <- na.omit(concatenated_data)

# Transform salary to PKR (1 USD = 285 PKR)
combined_data_clean <- combined_data_clean %>%
  mutate(salary_in_pkr = salary_in_usd * 285)

# Create summary table based on available columns
if("remote_ratio" %in% colnames(combined_data_clean)) {
  table <- combined_data_clean %>% 
    group_by(job_title) %>% 
    summarize(
      `Mean Salary (USD)` = round(mean(salary_in_usd, na.rm = TRUE), 2), 
      `SD Salary (USD)` = round(sd(salary_in_usd, na.rm = TRUE), 2),
      `Mean Remote Ratio` = round(mean(remote_ratio, na.rm = TRUE), 2),
      `SD Remote Ratio` = round(sd(remote_ratio, na.rm = TRUE), 2),
      `Employee Count` = n()
    ) %>%
    arrange(desc(`Mean Salary (USD)`))
} else {
  table <- combined_data_clean %>% 
    group_by(job_title) %>% 
    summarize(
      `Mean Salary (USD)` = round(mean(salary_in_usd, na.rm = TRUE), 2), 
      `SD Salary (USD)` = round(sd(salary_in_usd, na.rm = TRUE), 2),
      `Employee Count` = n()
    ) %>%
    arrange(desc(`Mean Salary (USD)`))
}
```

Overview {data-icon="fa-dashboard"}
=====================================

Column {data-width=250}
-----------------------------------------------------------------------

### Total Records {.value-box}

```{r}
renderValueBox({
  valueBox(
    value = formatC(nrow(combined_data_clean), format="d", big.mark=","),
    caption = "Total Salary Records",
    icon = "fa-database",
    color = "primary"
  )
})
```

### Average Salary (USD) {.value-box}

```{r}
renderValueBox({
  avg_salary <- mean(combined_data_clean$salary_in_usd, na.rm = TRUE)
  valueBox(
    value = paste0("$", formatC(round(avg_salary), format="d", big.mark=",")),
    caption = "Average Salary (USD)",
    icon = "fa-dollar-sign",
    color = "success"
  )
})
```

### Average Salary (PKR) {.value-box}

```{r}
renderValueBox({
  avg_salary_pkr <- mean(combined_data_clean$salary_in_pkr, na.rm = TRUE)
  valueBox(
    value = paste0("₨ ", formatC(round(avg_salary_pkr), format="d", big.mark=",")),
    caption = "Average Salary (PKR)",
    icon = "fa-money-bill-wave",
    color = "warning"
  )
})
```

### Unique Job Titles {.value-box}

```{r}
renderValueBox({
  valueBox(
    value = length(unique(combined_data_clean$job_title)),
    caption = "Different Job Roles",
    icon = "fa-briefcase",
    color = "info"
  )
})
```

Column {data-width=750}
-----------------------------------------------------------------------

### 📊 Salary Distribution Overview

```{r}
renderPlotly({
  p <- ggplot(combined_data_clean, aes(x = salary_in_usd)) +
    geom_histogram(bins = 50, fill = "#007bff", color = "white", alpha = 0.8) +
    labs(title = "Distribution of Salaries (USD)",
         x = "Salary (USD)",
         y = "Frequency") +
    theme_minimal(base_size = 14) +
    scale_x_continuous(labels = comma) +
    theme(plot.title = element_text(hjust = 0.5, face = "bold", color = "#007bff"))
  
  ggplotly(p) %>%
    layout(hovermode = "x unified")
})
```

---

**📌 Dataset Summary**

This dashboard analyzes salary data from **2020 to 2023**, covering multiple job roles across various company sizes and employee locations. The exchange rate used is **1 USD = 285 PKR**.

---

Salary Analysis by Job {data-icon="fa-briefcase"}
=========================================

Column {.sidebar data-width=250}
-----------------------------------------

<div class="chart-title">
### 📋 Job Salary Analysis
</div>

```{r}
HTML("<hr>")
h3(strong("Welcome!"))
h4("This is the Final Project Dashboard")
HTML("<hr>")
h3(strong(em("The table displays comprehensive salary statistics for various job roles:")))
h4("✓ Average Salary in USD")
h4("✓ Standard Deviation")
if("remote_ratio" %in% colnames(combined_data_clean)) {
  h4("✓ Remote Work Ratio")
}
h4("✓ Employee Count")
HTML("<hr>")
h4(em("💡 Use the search bar to find specific jobs"))
h4(em("📊 Click column headers to sort"))
h4(em("📥 Export data using the buttons above the table"))
```

Column {data-width=750}
-----------------------------------------

### 📊 Salary Statistics by Job Title

```{r}
renderDT({
  datatable(table, 
            extensions = 'Buttons',
            options = list(
              dom = 'Bfrtip',
              buttons = list(
                list(extend = 'copy', className = 'btn-primary'),
                list(extend = 'csv', className = 'btn-success'),
                list(extend = 'excel', className = 'btn-info', title = 'Salary_Analysis'),
                list(extend = 'pdf', className = 'btn-danger', title = 'Salary_Analysis'),
                list(extend = 'print', className = 'btn-warning')
              ),
              pageLength = 15,
              lengthMenu = c(10, 15, 25, 50, 100),
              scrollX = TRUE,
              scrollY = "500px",
              searchHighlight = TRUE,
              initComplete = JS(
                "function(settings, json) {",
                "$(this.api().table().header()).css({'background-color': '#007bff', 'color': '#fff'});",
                "}"
              )
            ),
            rownames = FALSE,
            class = 'cell-border stripe hover') %>%
    formatCurrency(c('Mean Salary (USD)', 'SD Salary (USD)'), '$', digits = 0) %>%
    formatStyle(
      'Mean Salary (USD)',
      background = styleColorBar(range(table$`Mean Salary (USD)`), '#e3f2fd'),
      backgroundSize = '100% 90%',
      backgroundRepeat = 'no-repeat',
      backgroundPosition = 'center'
    )
})
```

Company Size Analysis {data-icon="fa-building"}
=========================================

Column {.sidebar data-width=250}
-----------------------------------------

<div class="chart-title">
### 🏢 Company Size Impact
</div>

```{r}
HTML("<hr>")
h3(strong(em("Company Size Classification:")))
h4("🔵 L = Large Companies")
h4("🟢 M = Medium Companies")
h4("🟡 S = Small Companies")
HTML("<hr>")
h3(strong(em("The bar plot compares average salaries across different company sizes.")))
h4("📈 Hover over bars for detailed information")
h4("🔍 Interactive zoom and pan available")
HTML("<hr>")
h4(em("Key Insight: Company size often correlates with salary levels"))
```

Column {data-width=750}
-----------------------------------------

### 📊 Average Salary by Company Size (USD)

```{r}
renderPlotly({
  bar_plot_company_size <- combined_data_clean %>%
    group_by(company_size) %>%
    summarize(
      mean_salary = mean(salary_in_usd, na.rm = TRUE),
      count = n(),
      sd_salary = sd(salary_in_usd, na.rm = TRUE)
    ) %>%
    arrange(desc(mean_salary))
  
  colors <- c("L" = "#0066cc", "M" = "#00cc66", "S" = "#ffcc00")
  
  plot_ly(bar_plot_company_size, 
          x = ~company_size, 
          y = ~mean_salary,
          type = 'bar',
          color = ~company_size,
          colors = colors,
          text = ~paste0("Size: ", company_size, 
                        "<br>Avg Salary: $", formatC(round(mean_salary), format="d", big.mark=","),
                        "<br>Employees: ", count,
                        "<br>Std Dev: $", formatC(round(sd_salary), format="d", big.mark=",")),
          hoverinfo = 'text') %>%
    layout(
      title = list(
        text = "<b>Company Size vs. Average Salary</b>",
        font = list(size = 18, color = "#007bff")
      ),
      xaxis = list(
        title = "<b>Company Size</b>",
        titlefont = list(size = 14),
        tickfont = list(size = 12)
      ),
      yaxis = list(
        title = "<b>Average Salary (USD)</b>",
        titlefont = list(size = 14),
        tickfont = list(size = 12),
        tickformat = "$,.0f"
      ),
      showlegend = FALSE,
      plot_bgcolor = "#f8f9fa",
      paper_bgcolor = "#ffffff"
    )
})
```

---

### 💰 Salary Comparison: USD vs PKR

```{r}
renderPlotly({
  company_comparison <- combined_data_clean %>%
    group_by(company_size) %>%
    summarize(
      `USD` = mean(salary_in_usd, na.rm = TRUE),
      `PKR` = mean(salary_in_pkr, na.rm = TRUE) / 1000  # In thousands
    ) %>%
    pivot_longer(cols = c(USD, PKR), names_to = "Currency", values_to = "Salary")
  
  plot_ly(company_comparison,
          x = ~company_size,
          y = ~Salary,
          color = ~Currency,
          colors = c("#007bff", "#28a745"),
          type = 'bar',
          text = ~ifelse(Currency == "USD",
                        paste0("$", formatC(round(Salary), format="d", big.mark=",")),
                        paste0("₨ ", formatC(round(Salary*1000), format="d", big.mark=","))),
          textposition = 'outside',
          hovertemplate = paste0("<b>%{x}</b><br>",
                                "%{text}<br>",
                                "<extra></extra>")) %>%
    layout(
      title = list(text = "<b>Salary Comparison Across Company Sizes</b>"),
      xaxis = list(title = "<b>Company Size</b>"),
      yaxis = list(title = "<b>Average Salary</b>"),
      barmode = 'group',
      legend = list(title = list(text = "<b>Currency</b>"))
    )
})
```

Employee Residence {data-icon="fa-globe"}
=========================================

Column {.sidebar data-width=250}
-----------------------------------------

<div class="chart-title">
### 🌍 Global Salary Distribution
</div>

```{r}
HTML("<hr>")
h3(strong("Geographic Analysis"))
HTML("<hr>")
h3(strong(em("This visualization shows average salaries based on employee residence countries.")))
h4("🌎 Each bar represents a country")
h4("📊 Height indicates average salary")
h4("🔢 Hover for exact figures")
HTML("<hr>")
h4(em("💡 Top countries by salary are shown"))
h4(em("📍 Location impacts compensation significantly"))
HTML("<hr>")

selectInput("top_n_residence", 
            "Number of countries to display:",
            choices = c(10, 15, 20, 25, 30),
            selected = 15)
```

Column {data-width=750}
-----------------------------------------

### 🗺️ Average Salary by Employee Residence (Top Countries)

```{r}
renderPlotly({
  n_countries <- as.numeric(input$top_n_residence)
  
  bar_plot_residence <- combined_data_clean %>%
    group_by(employee_residence) %>%
    summarise(
      mean_salary = mean(salary_in_usd, na.rm = TRUE),
      count = n(),
      sd_salary = sd(salary_in_usd, na.rm = TRUE)
    ) %>%
    arrange(desc(mean_salary)) %>%
    head(n_countries)
  
  plot_ly(bar_plot_residence, 
          x = ~reorder(employee_residence, mean_salary), 
          y = ~mean_salary,
          type = 'bar',
          marker = list(
            color = ~mean_salary,
            colorscale = 'Viridis',
            showscale = TRUE,
            colorbar = list(title = "Salary (USD)")
          ),
          text = ~paste0("Country: ", employee_residence,
                        "<br>Avg Salary: $", formatC(round(mean_salary), format="d", big.mark=","),
                        "<br>Employees: ", count,
                        "<br>Std Dev: $", formatC(round(sd_salary), format="d", big.mark=",")),
          hoverinfo = 'text') %>%
    layout(
      title = list(
        text = paste0("<b>Top ", n_countries, " Countries by Average Salary</b>"),
        font = list(size = 18, color = "#007bff")
      ),
      xaxis = list(
        title = "<b>Employee Residence</b>",
        titlefont = list(size = 14),
        tickangle = -45
      ),
      yaxis = list(
        title = "<b>Average Salary (USD)</b>",
        titlefont = list(size = 14),
        tickformat = "$,.0f"
      ),
      margin = list(b = 120),
      plot_bgcolor = "#f8f9fa",
      paper_bgcolor = "#ffffff"
    )
})
```

---

### 📈 Salary Trends by Top 10 Residence Countries

```{r}
renderPlotly({
  top_10_countries <- combined_data_clean %>%
    group_by(employee_residence) %>%
    summarise(mean_salary = mean(salary_in_usd, na.rm = TRUE)) %>%
    arrange(desc(mean_salary)) %>%
    head(10) %>%
    pull(employee_residence)
  
  trend_data <- combined_data_clean %>%
    filter(employee_residence %in% top_10_countries) %>%
    group_by(employee_residence, work_year) %>%
    summarise(avg_salary = mean(salary_in_usd, na.rm = TRUE), .groups = 'drop')
  
  plot_ly(trend_data,
          x = ~work_year,
          y = ~avg_salary,
          color = ~employee_residence,
          type = 'scatter',
          mode = 'lines+markers',
          line = list(width = 3),
          marker = list(size = 8)) %>%
    layout(
      title = list(text = "<b>Salary Trends Over Years (Top 10 Countries)</b>"),
      xaxis = list(title = "<b>Year</b>"),
      yaxis = list(title = "<b>Average Salary (USD)</b>", tickformat = "$,.0f"),
      legend = list(title = list(text = "<b>Country</b>")),
      hovermode = "x unified"
    )
})
```

Data Explorer {data-icon="fa-table"}
=========================================

### 📋 Complete Dataset

```{r}
renderDT({
  display_data <- combined_data_clean %>%
    mutate(
      `Salary (USD)` = salary_in_usd,
      `Salary (PKR)` = salary_in_pkr
    )
  
  datatable(display_data,
            extensions = 'Buttons',
            options = list(
              dom = 'Bfrtip',
              buttons = list(
                list(extend = 'copy', className = 'btn-primary'),
                list(extend = 'csv', className = 'btn-success', filename = 'complete_salary_data'),
                list(extend = 'excel', className = 'btn-info', filename = 'complete_salary_data'),
                list(extend = 'pdf', className = 'btn-danger', filename = 'complete_salary_data'),
                list(extend = 'print', className = 'btn-warning')
              ),
              pageLength = 25,
              lengthMenu = c(10, 25, 50, 100, 500),
              scrollX = TRUE,
              scrollY = "600px",
              searchHighlight = TRUE,
              initComplete = JS(
                "function(settings, json) {",
                "$(this.api().table().header()).css({'background-color': '#007bff', 'color': '#fff', 'font-weight': 'bold'});",
                "}"
              )
            ),
            filter = 'top',
            rownames = FALSE,
            class = 'cell-border stripe hover compact') %>%
    formatCurrency('Salary (USD)', '$', digits = 0) %>%
    formatCurrency('Salary (PKR)', '₨ ', digits = 0)
})
```