---
title: "iAC Studio — Website Performance & SEO Audit"
subtitle: "Live data analysis of iac-studio.com"
author: "Ahmad Ammar — Data Analyst"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    theme: cosmo
    highlight: tango
    code_folding: hide
    df_print: paged
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.align = "center",
  fig.width = 9,
  fig.height = 5
)

library(tidyverse)
library(httr)
library(jsonlite)
library(rvest)
library(xml2)
library(knitr)
library(kableExtra)
library(scales)

# ---- CONFIG ----
TARGET_URL    <- "https://iac-studio.com/"
EXTRA_PAGES   <- c("https://iac-studio.com/our-services/")
USER_AGENT    <- "Mozilla/5.0 (Windows NT 10.0; Win64; x64) AppleWebKit/537.36 (KHTML, like Gecko) Chrome/124.0 Safari/537.36"
PSI_ENDPOINT  <- "https://www.googleapis.com/pagespeedonline/v5/runPagespeed"

`%||%` <- function(a, b) if (is.null(a) || length(a) == 0 || (length(a) == 1 && is.na(a))) b else a

style_table <- function(df, caption = NULL) {
  kable(df, caption = caption) %>%
    kable_styling(
      bootstrap_options = c("striped", "hover", "condensed", "responsive"),
      full_width = TRUE,
      position = "left"
    )
}

# Retry helper: attempts fn() up to `tries` times with exponential backoff.
# Returns NULL on ultimate failure instead of stopping the report.
with_retry <- function(fn, tries = 3, base_wait = 2) {
  for (i in seq_len(tries)) {
    out <- tryCatch(fn(), error = function(e) {
      message(sprintf("Attempt %d/%d failed: %s", i, tries, e$message))
      NULL
    })
    if (!is.null(out)) return(out)
    if (i < tries) Sys.sleep(base_wait * (2 ^ (i - 1)))
  }
  NULL
}

notice <- function(msg) {
  cat(sprintf(
    "<div style='padding:12px;border-left:4px solid #E67E22;background:#FFF3E0;color:#7A4F01;border-radius:4px;'>%s</div>",
    msg
  ))
}
```

# Executive Summary

This report is a **live, data-driven audit** of `r TARGET_URL`. Every metric below is fetched in real time when this report is generated — from Google's PageSpeed Insights API, from the live HTML of the site, and from the network response headers returned by the iAC Studio servers.

The goal: identify exactly where the site is losing visitors, search rankings, or load speed — and give iAC Studio a prioritized action list to fix it.

**Report generated on:** `r format(Sys.time(), '%Y-%m-%d %H:%M:%S %Z')`

---

# Methodology

| Data Source | What it tells us | Real-time? |
|---|---|---|
| Google PageSpeed Insights API | Performance score, Core Web Vitals, accessibility, SEO, best practices | Yes — fetched live |
| Live HTML scrape (`rvest`) | Title tags, meta descriptions, headings, links, images, alt text | Yes — fetched live |
| HTTP response headers | Server tech, security headers, caching, compression | Yes — fetched live |
| Site structure crawl | Internal vs external links, page weight | Yes — fetched live |

> **Note for the client:** This audit uses *publicly available* data only. A deeper engagement — with access to your Google Analytics, Google Search Console, and App Store / Play Console — would unlock visitor behavior, conversion funnels, keyword rankings and install attribution data, which can multiply the actionable insight in this report.

---

# 1. Live Page Performance (Google PageSpeed Insights)

```{r psi-fetch, error=TRUE}
fetch_psi <- function(url, strategy = "mobile") {
  q <- paste0(
    PSI_ENDPOINT,
    "?url=", URLencode(url, reserved = TRUE),
    "&strategy=", strategy,
    "&category=performance",
    "&category=seo",
    "&category=accessibility",
    "&category=best-practices"
  )
  with_retry(function() {
    resp <- GET(q, timeout(90))
    if (http_error(resp)) stop("HTTP ", status_code(resp))
    content(resp, as = "parsed", type = "application/json")
  })
}

psi_mobile  <- fetch_psi(TARGET_URL, "mobile")
psi_desktop <- fetch_psi(TARGET_URL, "desktop")

extract_scores <- function(psi, label) {
  if (is.null(psi)) return(NULL)
  cats <- psi$lighthouseResult$categories
  tibble(
    Strategy      = label,
    Performance   = round(cats$performance$score * 100),
    SEO           = round(cats$seo$score * 100),
    Accessibility = round(cats$accessibility$score * 100),
    `Best Practices` = round(cats$`best-practices`$score * 100)
  )
}

scores <- bind_rows(
  extract_scores(psi_mobile,  "Mobile"),
  extract_scores(psi_desktop, "Desktop")
)
```

```{r psi-table, results='asis'}
if (!is.null(scores) && nrow(scores) > 0) {
  print(style_table(scores, caption = "Lighthouse scores (0–100) — live from Google PageSpeed Insights"))
} else {
  notice("Google PageSpeed Insights API did not return data after 3 attempts. This usually means a temporary rate limit or network blip. Re-knit the report in 60 seconds, or set a free PSI API key in the setup chunk.")
}
```

```{r psi-chart, fig.height=4, error=TRUE}
if (!is.null(scores) && nrow(scores) > 0) {
  scores %>%
    pivot_longer(-Strategy, names_to = "Category", values_to = "Score") %>%
    ggplot(aes(x = Category, y = Score, fill = Strategy)) +
    geom_col(position = position_dodge(width = 0.8), width = 0.7) +
    geom_text(aes(label = Score),
              position = position_dodge(width = 0.8),
              vjust = -0.4, size = 4, fontface = "bold") +
    geom_hline(yintercept = 90, linetype = "dashed", color = "darkgreen") +
    geom_hline(yintercept = 50, linetype = "dashed", color = "orange") +
    scale_fill_manual(values = c("Mobile" = "#2E86AB", "Desktop" = "#A23B72")) +
    scale_y_continuous(limits = c(0, 110), breaks = seq(0, 100, 25)) +
    labs(
      title = "Lighthouse Scores: Mobile vs Desktop",
      subtitle = "Green line = 'Good' (≥90). Orange line = 'Needs Improvement' (≥50).",
      x = NULL, y = "Score (0–100)"
    ) +
    theme_minimal(base_size = 12) +
    theme(legend.position = "top",
          plot.title = element_text(face = "bold"))
}
```

## Core Web Vitals (Mobile)

These are the three metrics Google uses for ranking:

```{r core-web-vitals}
extract_cwv <- function(psi) {
  if (is.null(psi)) return(NULL)
  audits <- psi$lighthouseResult$audits
  tibble(
    Metric = c("Largest Contentful Paint (LCP)",
               "Total Blocking Time (TBT)",
               "Cumulative Layout Shift (CLS)",
               "First Contentful Paint (FCP)",
               "Speed Index"),
    Value = c(
      audits$`largest-contentful-paint`$displayValue %||% NA,
      audits$`total-blocking-time`$displayValue       %||% NA,
      audits$`cumulative-layout-shift`$displayValue   %||% NA,
      audits$`first-contentful-paint`$displayValue    %||% NA,
      audits$`speed-index`$displayValue               %||% NA
    ),
    Score = c(
      round((audits$`largest-contentful-paint`$score %||% NA) * 100),
      round((audits$`total-blocking-time`$score       %||% NA) * 100),
      round((audits$`cumulative-layout-shift`$score   %||% NA) * 100),
      round((audits$`first-contentful-paint`$score    %||% NA) * 100),
      round((audits$`speed-index`$score               %||% NA) * 100)
    )
  )
}

cwv <- extract_cwv(psi_mobile)
if (!is.null(cwv)) {
  style_table(cwv, caption = "Core Web Vitals (mobile) — live from PSI")
}
```

---

# 2. On-Page SEO Audit (Live HTML Scrape)

```{r scrape-page, error=TRUE}
fetch_html <- function(url) {
  with_retry(function() {
    resp <- GET(
      url,
      user_agent(USER_AGENT),
      add_headers(`Accept-Language` = "en-US,en;q=0.9",
                  Accept = "text/html,application/xhtml+xml,application/xml;q=0.9,*/*;q=0.8"),
      timeout(60)
    )
    if (http_error(resp)) stop("HTTP ", status_code(resp))
    body <- content(resp, as = "text", encoding = "UTF-8")
    if (is.null(body) || nchar(body) < 100) stop("Empty or truncated response")
    list(response = resp, html = read_html(body))
  })
}

page <- fetch_html(TARGET_URL)

if (is.null(page)) {
  notice(paste0(
    "Could not fetch <code>", TARGET_URL, "</code> after 3 retries. ",
    "Possible causes: site is temporarily down, blocking automated requests, ",
    "or your network is restricting outbound HTTPS. ",
    "Sections 2–5 below will be skipped. Re-knit when the site is reachable."
  ))
}
```

```{r seo-extract, error=TRUE}
seo_summary <- function(page, url) {
  if (is.null(page)) {
    return(tibble(Field = "Status", Value = "Could not fetch page"))
  }
  doc <- page$html
  title    <- doc %>% html_element("title") %>% html_text(trim = TRUE)
  meta_desc <- doc %>% html_element("meta[name='description']") %>% html_attr("content")
  meta_kw   <- doc %>% html_element("meta[name='keywords']") %>% html_attr("content")
  canonical <- doc %>% html_element("link[rel='canonical']") %>% html_attr("href")
  og_title  <- doc %>% html_element("meta[property='og:title']") %>% html_attr("content")
  og_image  <- doc %>% html_element("meta[property='og:image']") %>% html_attr("content")
  viewport  <- doc %>% html_element("meta[name='viewport']") %>% html_attr("content")
  robots    <- doc %>% html_element("meta[name='robots']") %>% html_attr("content")

  flag <- function(x, good_min = 1, good_max = Inf) {
    if (is.na(x) || x == "") return("Missing")
    n <- nchar(x)
    if (n < good_min) return(paste0("Too short (", n, " chars)"))
    if (n > good_max) return(paste0("Too long (", n, " chars)"))
    paste0("OK (", n, " chars)")
  }

  tibble(
    Field = c("URL", "Title Tag", "Title length status",
              "Meta Description", "Meta Description status",
              "Meta Keywords", "Canonical URL",
              "Open Graph Title", "Open Graph Image",
              "Viewport (mobile)", "Robots directive"),
    Value = c(
      url,
      title %||% "Missing",
      flag(title, 30, 60),
      meta_desc %||% "Missing",
      flag(meta_desc, 70, 160),
      meta_kw %||% "Missing",
      canonical %||% "Missing",
      og_title %||% "Missing",
      og_image %||% "Missing",
      viewport %||% "Missing",
      robots %||% "Not set (defaults to index, follow)"
    )
  )
}

seo_tbl <- seo_summary(page, TARGET_URL)
style_table(seo_tbl, caption = "On-page SEO snapshot (live from iac-studio.com)")
```

```{r seo-extract-fallback, results='asis', echo=FALSE}
if (is.null(page)) notice("Skipping on-page SEO checks because the page could not be fetched.")
```

## Heading Structure (H1 → H6)

```{r headings, error=TRUE}
heading_audit <- function(page) {
  if (is.null(page)) return(tibble())
  map_dfr(1:6, function(i) {
    nodes <- page$html %>% html_elements(paste0("h", i)) %>% html_text(trim = TRUE)
    tibble(Level = paste0("H", i), Count = length(nodes),
           Examples = paste(head(nodes, 3), collapse = " | "))
  })
}

h_tbl <- heading_audit(page)
if (nrow(h_tbl) > 0) {
  style_table(h_tbl, caption = "Heading distribution and examples")

  ggplot(h_tbl, aes(x = Level, y = Count, fill = Level)) +
    geom_col() +
    geom_text(aes(label = Count), vjust = -0.4, fontface = "bold") +
    scale_fill_brewer(palette = "Blues") +
    labs(title = "Heading Tag Distribution", x = NULL, y = "Count") +
    theme_minimal(base_size = 12) +
    theme(legend.position = "none")
}
```

> **SEO rule of thumb:** A page should have **exactly one H1**, a logical descending hierarchy (don't skip from H1 to H4), and headings that include relevant keywords.

---

# 3. Content & Link Profile

```{r content-stats, error=TRUE, results='asis'}
content_stats <- function(page, url) {
  if (is.null(page)) return(NULL)
  doc <- page$html
  body_node <- doc %>% html_element("body")
  body_text <- if (inherits(body_node, "xml_missing")) "" else html_text(body_node, trim = TRUE)
  words <- if (nchar(body_text) > 0) str_split(body_text, "\\s+")[[1]] else character(0)
  words <- words[nchar(words) > 0]

  links <- doc %>% html_elements("a") %>% html_attr("href")
  links <- links[!is.na(links) & links != "" & !str_starts(links, "#") & !str_starts(links, "javascript:")]
  domain <- httr::parse_url(url)$hostname
  is_internal <- if (length(links) > 0) {
    str_detect(links, fixed(domain)) | str_starts(links, "/")
  } else logical(0)

  imgs <- doc %>% html_elements("img")
  alt_texts <- imgs %>% html_attr("alt")
  imgs_missing_alt <- sum(is.na(alt_texts) | alt_texts == "")

  tibble(
    Metric = c("Total word count",
               "Total links",
               "Internal links",
               "External links",
               "Total images",
               "Images missing alt text",
               "Alt-text coverage"),
    Value = c(
      comma(length(words)),
      comma(length(links)),
      comma(sum(is_internal)),
      comma(sum(!is_internal)),
      comma(length(imgs)),
      comma(imgs_missing_alt),
      ifelse(length(imgs) > 0,
             percent(1 - imgs_missing_alt / length(imgs), accuracy = 0.1),
             "n/a")
    )
  )
}

cs <- content_stats(page, TARGET_URL)
if (!is.null(cs)) {
  print(style_table(cs, caption = "Content & link profile (live)"))
} else {
  notice("Skipping content & link profile because the page could not be fetched.")
}
```

---

# 4. Server, Security & Tech Stack

```{r security-headers, error=TRUE, results='asis'}
security_audit <- function(page) {
  if (is.null(page)) return(NULL)
  h <- headers(page$response)
  check <- function(name, recommended = "Set any value") {
    val <- h[[tolower(name)]]
    tibble(
      Header = name,
      Status = ifelse(is.null(val), "MISSING", "PRESENT"),
      Value  = ifelse(is.null(val), paste("⚠️", recommended), substr(val, 1, 80))
    )
  }

  bind_rows(
    check("Strict-Transport-Security", "Add HSTS to enforce HTTPS"),
    check("Content-Security-Policy",   "Add CSP to prevent XSS"),
    check("X-Frame-Options",           "Add to prevent clickjacking"),
    check("X-Content-Type-Options",    "Add 'nosniff' to prevent MIME-type sniffing"),
    check("Referrer-Policy",           "Recommend 'strict-origin-when-cross-origin'"),
    check("Permissions-Policy",        "Restrict browser features"),
    check("Server",                    "Server software"),
    check("Content-Encoding",          "gzip/br compression saves bandwidth")
  )
}

sec <- security_audit(page)
if (!is.null(sec)) {
  print(style_table(sec, caption = "Security & server response headers"))
} else {
  notice("Skipping security header check because the page could not be fetched.")
}
```

```{r tech-detect, error=TRUE, results='asis'}
detect_tech <- function(page) {
  if (is.null(page)) return(NULL)
  doc <- page$html
  raw <- as.character(doc)

  signals <- c(
    "WordPress"   = "wp-content|wp-includes|/wp-json/",
    "WooCommerce" = "woocommerce",
    "Elementor"   = "elementor",
    "jQuery"      = "jquery",
    "Bootstrap"   = "bootstrap",
    "Google Analytics 4" = "gtag\\(|G-[A-Z0-9]{8,}",
    "Google Tag Manager" = "googletagmanager.com/gtm.js",
    "Facebook Pixel"     = "connect.facebook.net.*fbevents",
    "Cloudflare"         = "cloudflare",
    "reCAPTCHA"          = "google.com/recaptcha"
  )

  tibble(
    Technology = names(signals),
    Detected = sapply(signals, function(p) ifelse(str_detect(raw, regex(p, ignore_case = TRUE)), "Yes", "No"))
  )
}

tech <- detect_tech(page)
if (!is.null(tech)) {
  print(style_table(tech, caption = "Technology fingerprint (best-effort detection)"))
} else {
  notice("Skipping tech fingerprint because the page could not be fetched.")
}
```

---

# 5. Multi-Page Comparison

```{r multipage, error=TRUE}
all_pages <- c(TARGET_URL, EXTRA_PAGES)

page_metrics <- map_dfr(all_pages, function(u) {
  p <- fetch_html(u)
  if (is.null(p)) {
    return(tibble(URL = u, Title = "FETCH FAILED", `Title len` = NA,
                  `Meta desc len` = NA, `H1 count` = NA, Words = NA, Links = NA, Images = NA))
  }
  doc <- p$html
  title <- doc %>% html_element("title") %>% html_text(trim = TRUE)
  desc  <- doc %>% html_element("meta[name='description']") %>% html_attr("content")
  h1s   <- doc %>% html_elements("h1") %>% length()
  body  <- doc %>% html_element("body") %>% html_text(trim = TRUE)
  words <- length(str_split(body, "\\s+")[[1]])
  links <- doc %>% html_elements("a") %>% length()
  imgs  <- doc %>% html_elements("img") %>% length()

  tibble(
    URL           = u,
    Title         = ifelse(is.na(title), "Missing", substr(title, 1, 50)),
    `Title len`   = nchar(title %||% ""),
    `Meta desc len` = nchar(desc %||% ""),
    `H1 count`    = h1s,
    Words         = words,
    Links         = links,
    Images        = imgs
  )
})

style_table(page_metrics, caption = "Page-by-page snapshot (live)")
```

---

# 6. Prioritized Recommendations

```{r recommendations, error=TRUE}
recommendations <- function(scores, cwv, seo_tbl, sec, cs) {
  recs <- list()

  # Performance
  if (!is.null(scores) && nrow(scores) > 0) {
    mob_perf <- scores$Performance[scores$Strategy == "Mobile"]
    if (length(mob_perf) > 0 && !is.na(mob_perf) && mob_perf < 90) {
      sev <- if (mob_perf < 50) "HIGH" else "MEDIUM"
      recs[[length(recs) + 1]] <- tibble(
        Priority = sev, Area = "Performance",
        Finding = paste0("Mobile performance score is ", mob_perf, "/100."),
        Action = "Compress and lazy-load images, defer non-critical JS, enable browser caching, use a CDN."
      )
    }
  }

  if (is.null(seo_tbl) || !"Field" %in% names(seo_tbl)) {
    return(tibble(Priority = "INFO", Area = "Data",
                  Finding = "Page could not be fetched, so on-page recommendations were skipped.",
                  Action = "Re-knit the report when the site is reachable."))
  }

  # Meta description
  meta_status <- seo_tbl$Value[seo_tbl$Field == "Meta Description status"]
  if (length(meta_status) > 0 && (str_detect(meta_status, "Missing|Too"))) {
    recs[[length(recs) + 1]] <- tibble(
      Priority = "HIGH", Area = "SEO",
      Finding = paste("Meta description issue:", meta_status),
      Action = "Write a compelling 120–155 character meta description with primary keyword + value proposition."
    )
  }

  # Title tag
  title_status <- seo_tbl$Value[seo_tbl$Field == "Title length status"]
  if (length(title_status) > 0 && str_detect(title_status, "Missing|Too")) {
    recs[[length(recs) + 1]] <- tibble(
      Priority = "HIGH", Area = "SEO",
      Finding = paste("Title tag issue:", title_status),
      Action = "Rewrite title to 50–60 characters including main service keyword (e.g. 'Mobile App Development')."
    )
  }

  # Security headers
  if (!is.null(sec)) {
    missing_sec <- sec %>% filter(Status == "MISSING", Header != "Server")
    if (nrow(missing_sec) > 0) {
      recs[[length(recs) + 1]] <- tibble(
        Priority = "MEDIUM", Area = "Security",
        Finding = paste(nrow(missing_sec), "security headers missing:",
                        paste(missing_sec$Header, collapse = ", ")),
        Action = "Configure security headers via .htaccess or web server config to improve trust signals & ranking."
      )
    }
  }

  # Image alt text
  if (!is.null(cs)) {
    alt_row <- cs$Value[cs$Metric == "Images missing alt text"]
    if (length(alt_row) > 0 && as.numeric(gsub(",", "", alt_row)) > 0) {
      recs[[length(recs) + 1]] <- tibble(
        Priority = "MEDIUM", Area = "Accessibility / SEO",
        Finding = paste(alt_row, "images are missing alt text."),
        Action = "Add descriptive alt text to every <img> for accessibility and image-search rankings."
      )
    }
  }

  if (length(recs) == 0) {
    return(tibble(Priority = "INFO", Area = "—",
                  Finding = "No critical issues detected from public data.",
                  Action = "Request analytics access for deeper behavioural analysis."))
  }
  bind_rows(recs) %>% arrange(factor(Priority, levels = c("HIGH", "MEDIUM", "LOW", "INFO")))
}

rec_tbl <- recommendations(scores, cwv, seo_tbl, sec, cs)

rec_tbl %>%
  kable(caption = "Prioritized action list") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = TRUE) %>%
  column_spec(1, bold = TRUE,
              color = "white",
              background = case_when(
                rec_tbl$Priority == "HIGH"   ~ "#C0392B",
                rec_tbl$Priority == "MEDIUM" ~ "#E67E22",
                rec_tbl$Priority == "LOW"    ~ "#27AE60",
                TRUE                          ~ "#2980B9"
              ))
```

---

# 7. What's Next — Unlocking Real User Data

This audit measured what is **publicly visible**. The next layer of analysis — the one that drives revenue decisions — needs access to your private data. With the following granted, the next report will include:

| If you grant access to… | I can analyze… |
|---|---|
| **Google Analytics 4** | Real visitor counts, traffic sources, bounce rate, session duration, conversion funnels |
| **Google Search Console** | Actual keyword rankings, click-through rates, impressions, indexing issues |
| **Google Play Console / App Store Connect** | Install rate, retention curves, crash analytics, store impression-to-install funnel |
| **Server / hosting access** | Log-level analysis of bots, security threats, slow endpoints |

---

```{r session-info, echo=FALSE}
cat("Report rendered with R", R.version.string, "on", format(Sys.time(), "%Y-%m-%d %H:%M:%S %Z"))
```

---

*Prepared by **Ahmad Ammar** — Data Analyst*
